---
title: "gcimputeR Vignette on low rank Gaussian copula with quantified uncertainty"
author: "Yuxuan Zhao"
date: "9/1/2020"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{gcimputeR Vignette on low rank Gaussian copula with quantified uncertainty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newcommand{\bX}{\mathbf{X}}
\newcommand{\bZ}{\mathbf{Z}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bz}{\mathbf{z}}
\newcommand{\bigf}{\mathbf{f}}
\newcommand{\bo}{\mathbf{0}}


## Generate data 

We generate the high rank continuous data matrix $\bX\in \mathbb{R}^{500\times200}$ from a low rank Gaussian copula as in the experiments of [our paper](https://proceedings.neurips.cc/paper/2020/hash/f076073b2082f8741a9cd07b789c77a0-Abstract.html) and randomly mask $40\%$ observation as test set.

```{r generate data}
  library(gcimputeR)
  set.seed(410)
  var_types = list('cont'=1:200)
  f = function(x)x^3
  X = generate_LRGC(rank = 10, sigma = 0.1,n = 500,  var_types = var_types, cont_transform = f)
  # mask 40% of the original observation
  X_obs = mask_MCAR(X, mask_fraction = 0.4)
```

## Fitting low rank Gaussian copula

Simply specify the rank to the function call. 

```{r fitting}
est = impute_LRGC(X_obs, rank = 10) # around 8 secs
print("Normalized root mean squared error (NRMSE) is: ")
print(round(cal_rmse(xhat = est$Ximp, xobs = X_obs, xtrue = X, relative = TRUE), 4))
```

## Construct confidence interval

```{r confidence interval}
ct = ct_impute(X_obs, est, 0.95)
loc = is.na(X_obs)
print("The empirical coverage is: ")
print(mean(X[loc] >= ct$lower[loc] & X[loc] <= ct$upper[loc]))
print("The mean confidence interval length is: ")
mean(ct$upper[loc] - ct$lower[loc])
```
